`---
title: "Covid19 - Data analysis"
output: html_notebook
---

Data source: https://github.com/CSSEGISandData/COVID-19/tree/master/csse_covid_19_data/csse_covid_19_time_series

```{r}
library(dplyr)
library(readr)
library(janitor)
library(ggplot2)
library(tidyr)
library(grid)
```


```{r}
#CONFIRM CASES FILE
confirmed <- read_csv('time_series_covid19_confirmed_global.csv')
head(confirmed)
```



```{r}
#Total case by 9th Nov
total_case <- confirmed %>% 
  summarise(sum(`11/9/20`))
  
total_case
```
```{r}
#Group by countries
group_countries <- confirmed%>%
    group_by(`Country/Region`) %>%
    summarise_at(vars(-`Province/State`, -Lat, -Long), sum)
group_countries
```
```{r}
#Total cases of countries from the north
north_equator <- confirmed %>%
    filter(Lat > 0) %>%
    summarise(sum(`11/9/20`))
north_equator

north_confirmed_percent <- north_equator*100/total_case
north_confirmed_percent

```

```{r}
#Total cases of countries from the south
south_equator <- confirmed %>%
    filter(Lat < 0) %>%
    summarise(sum(`11/9/20`))
south_equator

south_confirmed_percent <- south_equator*100/total_case
south_confirmed_percent
```

```{r}
#Chart to compare north and south
regions <- c('Northern countries', 'Southern countries')
total_cases <- c(40720033, 10192684)
percentage<- c(79.97893, 20.01963)

north_south_data <- data.frame(regions, total_cases, percentage)
north_south_data 


```
```{r}
#find the countries with maximum number of confirmed cases
just_number <- group_countries %>%
    select(-`Country/Region`) 
max_case <- max(just_number, na.rm=TRUE)
max_case

which(group_countries == max_case, arr.ind=T)
group_countries[182, 294]

#CONCLUSION: The country that experienced highest no. of cases is US. 10110552 cases on 9th Nov 2020
```
```{r}
#Top six countries 
top_six <- tail(group_countries[order(group_countries$`11/9/20`),]) 

top_six_table <- top_six %>%
    select(`Country/Region`,`11/9/20`) 
top_six_table



top_six_transposed <- top_six %>%
  t() %>%
  as.data.frame() %>%
  row_to_names(row_number = 1)


top_six_transposed <- top_six_transposed %>%
  apply(MARGIN=2, as.numeric) %>%
  as.data.frame()

top_six_transposed <- top_six_transposed %>%
  mutate(date = c(1: nrow(top_six_transposed)))

top_six_transposed
df <- gather(top_six_transposed, key = Country, value = cases, c("Spain", "Russia", "France", "Brazil", "India", "US"))

top_six_trend <- ggplot(df, aes(x=date, y = cases, group = Country, colour = Country)) + geom_line()+ labs(title='Confirmed cases in top six countries',x='From 22nd Jan to 9th Nov 2020', y='Number of cases')+ 
  scale_y_continuous(limit = c(0, 11000000)) +
  theme(axis.text.x = element_blank())

top_six_trend
ggsave('top_six_trend.png')
```
```{r}
#RECOVERED & DEATHS
recovered <- read_csv('time_series_covid19_recovered_global.csv')
head(recovered)
deaths <- read_csv('time_series_covid19_deaths_global.csv')
head(deaths)
      
```
```{r}
#Dropping the 2 last cols (11/10/20 & 11/11/20) to keep the dates consistent
recovered <- recovered %>% select (-`11/10/20`, -`11/11/20`)
deaths <- deaths %>% select(-`11/10/20`, -`11/11/20`)

head(deaths)


```


```{r}
#Total cases recovered by 9th Nov
sum_recovered <- recovered %>%
    summarise(sum(`11/9/20`))
sum_recovered
sum_recovered_percent <- sum_recovered*100/total_case
sum_recovered_percent

#Northern countries (Lat >0)
north_recovered <- recovered %>%
    filter(Lat>0) %>%
    summarise(sum(`11/9/20`))
north_recovered_percent <- north_recovered*100/north_equator
north_recovered_percent

#Southern countries (Lat<0)
south_recovered <- recovered %>%
    filter(Lat<0) %>%
    summarise(sum(`11/9/20`))
south_recovered_percent <- south_recovered*100/ south_equator
south_recovered_percent

#Total deaths by 9th Nov
sum_deaths <- deaths %>%
    summarise(sum(`11/9/20`))
sum_deaths

north_deaths <- deaths %>%
    filter(Lat>0) %>%
    summarise(sum(`11/9/20`))
north_deaths
north_deaths_percent <- north_deaths*100/north_equator
north_deaths_percent

south_deaths <- deaths %>%
    filter(Lat<0) %>%
    summarise(sum(`11/9/20`))
south_deaths
south_deaths_percent <- south_deaths*100/ south_equator
south_deaths_percent


#CONCLUSION! South recovered faster than North
```


```{r}

south_treatment <- south_equator - south_deaths - south_recovered
north_treatment <- north_equator - north_deaths - north_recovered

south_treatment
north_treatment

#comparison North and South table
regions <- c('Northern countries', 'Southern countries')
confirmed <- c(40720033, 10192684)
deaths <- c(955048, 308243)
on_treatment <- c(15634207, 726474)
recovered <- c(24130778, 9157967)

north_south_summary <- data.frame(regions, confirmed, deaths, on_treatment, recovered)
north_south_summary
```


```{r}
#Chart for 2 areas:
status <- c('Recovered', 'Dead', 'On treatment')
northern_countries <- c(24130778, 955048, 15634207)
southern_countries <- c(9157967, 308243, 726474)
north_and_south <- data.frame(status, northern_countries, southern_countries)

north_and_south

df <- pivot_longer(north_and_south, cols=-1)
df

percent_chart <- ggplot(df, aes(x=name, y=value, fill=status)) + geom_bar(position="fill", stat="identity") + scale_y_continuous(labels = scales::percent) + scale_fill_brewer() + theme_light() + theme(axis.title=element_blank())
percent_chart
ggsave('north_south_percent_chart.png')

north_and_south_chart <- ggplot(df, aes(x=name, y=value, fill=status)) + geom_bar(position="stack", stat="identity") + ylab('Number of cases') + scale_fill_brewer() + theme_classic() +  theme(axis.title.x=element_blank())
north_and_south_chart 
ggsave('north_and_south_chart.png')

```



```{r}




```
 
Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file   containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
t
The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
`
